---
title: "Subset cerebellum endothelial cells"
author: "Bono"
date: "18/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
load_lib <- "Load Libraries"
```
## `r load_lib`
* Load all the necessary libraries


```{r load_lib, message=FALSE}
library(Seurat) 
library(dplyr)
library(here)
library(ggplot2)
library(msigdbr)
library(clusterProfiler)
library(GSEABase)
library(org.Hs.eg.db)
library(SCINA)
library(tidyr)
library(scSorter)
```

```{r load_data, echo=TRUE}
full_matrix <- readRDS(here("data/srt_anno_01.RDS")) #Load entire dataset
Idents(full_matrix) <- "clusters_named" #Set identity to clusters_named column on metadata
```


```{r, include=FALSE}
subset_data <- "Subset CB Endothelial Cell Populations"
```
## `r subset_data`
* Subset for endothelial cells and pericytes populations
* From the EC and pericyte populations, further subset for CB specific endothelial-pericyte populations
* Re-calculate mitochondrial gene percentage. Since previously mito percentage is calculated separately for 
spliced and unspliced matrices, but only the spliced matrix counts mito genes. We expect recalculated mito percentage to be lower.


#Subset only CB cells labelled endo-peri1 & 2
```{r subset_data, echo=FALSE}
ecpc <- subset(full_matrix, ident = c("Endothelial-Pericyte_1", "Endothelial-Pericyte_2")) #Subset all EC
##Re-calculate mitochondrial gene percentage###
mt.genes <- rownames(ecpc)[grep("^MT-",rownames(ecpc))] #identify all mitochondrial genes
percent.mito <- PercentageFeatureSet(ecpc, pattern = "^MT-")
ecpc <- AddMetaData(ecpc, percent.mito, col.name = "percent.mito") 

###Subset for CB###
CB_EC <- subset(x = ecpc, subset=Tissue=="CB") #Subset BECs
Idents(CB_EC) <- "BECs" #Set identity classes
Idents(CB_EC)
tissue_spec <- rep("BEC", ncol(CB_EC)) 
CB_EC <- AddMetaData(CB_EC, tissue_spec, col.name = "Tissues") #Re-define
```



```{r, include=FALSE}
cluster_EC_Peri <- "Distinguish Endothelial Cells and Pericytes in CB Endo-Pericyte Population"
```
## `r cluster_EC_Peri`
Aims to distinguish between endothelial cells and pericytes of BEC (CB), through feature selection, dimension reduction, and clustering analysis.

* HVG = 2000
* Perform PCA
* PC used = 20
* Clustering


```{r cluster_EC_Peri, message=FALSE}
#Calculate based on variance of log-expression of genes across cells. 2000 HVGs are selected.
CB_EC <- FindVariableFeatures(CB_EC, selection.method = "vst", nfeatures = 2000)  

#Scale data for PCA
all.genes <- rownames(CB_EC)
CB_EC <- ScaleData(CB_EC, features = all.genes) 

#Run PCA
CB_EC <- RunPCA(CB_EC, features = VariableFeatures(object=CB_EC))
```


#Visualize
```{r}
VizDimLoadings(CB_EC, dims = 1:2, reduction = "pca") 

Idents(CB_EC) <- "clusters_named"
DimPlot(CB_EC, reduction = "pca")

ElbowPlot(CB_EC, ndims = 50)
```


```{r, message=FALSE, warning= FALSE}
CB_EC <- FindNeighbors(CB_EC, dims = 1:30) #KNN
CB_EC <- FindClusters(CB_EC, resolution = 0.1) #Louvain algorithm
Idents(CB_EC) <- "RNA_snn_res.0.1"

CB_EC <- RunUMAP(CB_EC, dims = 1:30)

DimPlot(CB_EC, reduction = "umap", label = T)
```


#Subset for ECs in CB population
The following markers were used to distinguish cell types:

* endothelial cells: VWF, FLT1, CLDN5, ABCB1, THSD4, PECAM1, CDH5
* pericytes: PDGFRB, NOTCH3, PRKG1, CSPG4 (NG2)

```{r, fig.width=15, fig.height=13}

#EC  Biomarkers
FeaturePlot(CB_EC, features = c("VWF",
                              "FLT1",
                              "CLDN5",
                              "ABCB1",
                              "ST6GALNAC3",
                              "ELOVL7",
                              "THSD4",
                              "PECAM1",
                              "MECOM"
                              ))
```


```{r, fig.width= 7, fig.height=10}
VlnPlot(CB_EC, features = c("VWF", "FLT1", "CLDN5", "ABCB1",  "ST6GALNAC3", "ELOVL7", "THSD4", "PECAM1", "MECOM"))
```



```{r, fig.width=15, fig.height=12}
#Pericyte Biomarkers
FeaturePlot(CB_EC, features = c("PDGFRB",
                              "NOTCH3",
                              "PRKG1",
                              "CSPG4",
                              "GPC5",
                              "RGS5",
                              "PDE7B"
                              ))
```


```{r, fig.width= 7, fig.height=10}
VlnPlot(CB_EC, features = c("PDGFRB", "NOTCH3", "PRKG1", "CSPG4",  "GPC5", "RGS5", "PDE7B"))
```


```{r}
CB_EC$annotate <- CB_EC$RNA_snn_res.0.1

Idents(CB_EC) <- "annotate"

CB_EC <- RenameIdents(CB_EC, "0" = "Endothelial", "1" = "Pericyte", "2" = "Endothelial")

CB_EC$annotate <- Idents(CB_EC)

DimPlot(CB_EC, label = T, repel = T, group.by = "annotate")
```


```{r, message= FALSE}
CB_EC <- subset(CB_EC, subset=annotate=="Endothelial")
saveRDS(CB_EC, "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/CB_annotated.RDS")
```




#############################################################################################



```{r, include=FALSE}
annot_CB <- "CB endothelial subtype annotation"
```


#Run feature selection
```{r, fig.width = 15, fig.height= 8, message= FALSE, warning=FALSE}
CB_EC <- FindVariableFeatures(CB_EC, selection.method = "vst", nfeatures= 2000)
```


#Run PCA
```{r}
all.genes <- rownames(CB_EC)

CB_EC <- ScaleData(CB_EC, features = all.genes) #scale all variable genes

CB_EC <- RunPCA(CB_EC, features = VariableFeatures(object=CB_EC))

ElbowPlot(CB_EC, ndims = 50)
```


#Cluster CB-ECs
```{r, message=FALSE, warning= FALSE}
CB_EC <- FindNeighbors(CB_EC, dims = 1:30) #KNN

CB_EC <- FindClusters(CB_EC, resolution = 0.2) #Louvain algorithm

Idents(CB_EC) <- "RNA_snn_res.0.2"

CB_EC <- RunUMAP(CB_EC, dims = 1:30)

DimPlot(CB_EC, reduction = "umap", label=T, repel=T)
```


```{r, include=FALSE}
feature_plt_1 <- "Feature Plots (Artery genes)"
```
# `r feature_plt_1`


```{r feature_plt_1, fig.width=15, fig.height=12}
FeaturePlot(CB_EC, features = c("BMX",     #According to Vanlandewijck paper (Mouse)
                                 "SEMA3G", #According to Vanlandewijck paper (Mouse)
                                 "VEGFC",  #According to Vanlandewijck paper (Mouse)
                                 "EFNB2",  #Preferentially on arterial endothelium (Human & mouse)
                                 "NOTCH1", #Arterial specificaiton and angiogenic potential (Mouse)
                                 "NOTCH4", 
                                 "DLL4",   #Canonical NOTCH ligand strongly expressed in arterial ECs
                                 "HEY2",    #Expressed in humans
                                 "MGP"
                                 ))
```


```{r, include=FALSE}
feature_plt_3 <- "Feature Plots (Capillary genes)"
```
# `r feature_plt_3`


```{r feature_plt_3, fig.width=15, fig.height=12}

FeaturePlot(CB_EC, features = c("SLC16A1",  #According to Vanlandewijck paper (Mouse)
                                 "MFSD2A",   #According to Vanlandewijck paper (Mouse)
                                 "SLC7A5",   #According to Vanlandewijck paper (Mouse)
                                 "TFRC",     #According to Vanlandewijck paper (Mouse)
                                 "ABCB1",    #P-glycoprotein (MDR) is often expressed in brain capillary ECs (ABC transporter)(Mouse)
                                 "ABCG2",    #BCRP also demonstrated before is excessively expressed in brain capillary ECs (ABC transporter)(Human & Mouse)
                                 "SLC7A5",   #LAT-1 expressed in BBB models, responsible for transport of large neutral AAs, L-dopa and gabapentin by brain capillary ECs (Human & Mouse)
                                 "SLC2A1"   #GLUT-1 expressed in BBB models, responsible for drug transporter in the brain capillary (Human & Mouse)
                                 ))
```


```{r, include=FALSE}
feature_plt_2 <- "Feature Plots (Vein genes)"
```
# `r feature_plt_2`


```{r feature_plt_2, fig.width=15, fig.height=12}
FeaturePlot(CB_EC, features = c("NR2F2",   #Expressed in vein ECs (Mouse)
                                 "VCAM1",   #According to Vanlandewijck paper (Mouse)
                                 "VWF",     #According to Vanlandewijck paper (Mouse)
                                 "SLC38A5", #According to Vanlandewijck paper (Mouse)
                                 "EPHB4", #Preferentially expressed in vein ECs (Human & mouse),
                                 "EMCN",  #Expressed in humans
                                 "IL1R1",  #Expressed in Humans
                                 "ICAM1"
                                 ))
```


```{r}
DimPlot(CB_EC, group.by = "Age")

DimPlot(CB_EC, group.by = "BBN")

DimPlot(CB_EC, group.by = "CauseOfDeath_category")
```


```{r, include=FALSE}
feature_plt_5 <- "Feature Plots (Based on cell quality)"
```
#`r feature_plt_5`


Clusters 0 and 1 might be clustered based on quality
```{r}
FeaturePlot(CB_EC, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"))
```


#Add Subtype Annotations
```{r, include=FALSE}
Annot <- "EC Subtype Annotations"
```


Perform scSorter (Cell assign approach 2)
```{r Annot, message= FALSE}
Type <- c("Capillary", "Capillary", "Capillary", "Capillary", "Arterial", "Arterial", "Arterial", "Arterial", "Vein", "Vein", "Vein", "Vein", "Vein")

Marker <- c("MFSD2A", "SLC16A1", "SLC7A5", "TFRC", "VEGFC", "HEY2","SEMA3G", "EFNB2", "ICAM1", "IL1R1","EPHB4", "NR2F2", "VCAM1")

Weight <- c(3, 3, 3, 3, 3, 5, 3, 3, 5, 5, 5, 5, 5)

Markerr <- cbind(Type, Marker, Weight)

Markerss <- as.data.frame(Markerr)
```


```{r}
Annot_EC_3 <- scSorter(CB_EC@assays$RNA@data, Markerss, alpha = 0)
```


```{r}
Idents(CB_EC) <- Annot_EC_3$Pred_Type

CB_EC$subtype <- Annot_EC_3$Pred_Type

CB_EC <- AddMetaData(CB_EC, metadata = Annot_EC_3$Pred_Type, col.name = "subtype")

CB_EC <- subset(CB_EC, subset = (subtype == "Arterial" | subtype == "Capillary" | subtype == "Vein"))

CB_EC$subtype <- Idents(CB_EC)

DimPlot(CB_EC, reduction = "umap", group.by = "subtype", cols = c("chartreuse4", "coral2", "cornflowerblue")) + ggtitle("CB (Cerebellum) endothelial subtype")
```


```{r, echo = TRUE}
CellsPerSubtype <- as.data.frame(tapply(
    CB_EC@meta.data$Barcode, CB_EC@meta.data$subtype, function(x) length(x) ))

names(CellsPerSubtype) <- "NumberOfCells"

CellsPerSubtype$Subtype <- rownames(CellsPerSubtype)

CellsPerSubtype$Subtype <- rownames(CellsPerSubtype)

CellsPerSubtype
```


```{r}
art <- CB_EC[,CB_EC@meta.data$subtype == "Arterial"]

capi <- CB_EC[,CB_EC@meta.data$subtype == "Capillary"]

veins <- CB_EC[,CB_EC@meta.data$subtype == "Vein"]

tgt <- merge(art, c(capi, veins), add.cell.ids = c("Arterial", "Capillary", "Vein"))
```


```{r}
col_art <- rep("coral2",ncol(tgt[,tgt@meta.data$subtype=="Arterial"]))

col_capi <- rep("chartreuse4",ncol(tgt[,tgt@meta.data$subtype=="Capillary"]))

col_vein <- rep("cornflowerblue",ncol(tgt[,tgt@meta.data$subtype=="Vein"]))

palette <- c(col_art, col_capi, col_vein)

tgt$color <- palette
```


```{r, fig.width= 8, fig.height= 10}
par(mfrow = c(3, 1))
barplot(tgt@assays$RNA@data["HEY2",], main = "HEY2", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["VEGFC",], main = "VEGFC", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["MGP",], main = "MGP", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
#barplot(tgt@assays$RNA@data["EFNB2",], main = "EFNB2", axisnames = FALSE, ylab = "count", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 2)
barplot(tgt@assays$RNA@data["MFSD2A",], main = "MFSD2A", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["SLC16A1",], main = "SLC16A1", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["TFRC",], main = "TFRC", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
#barplot(tgt@assays$RNA@data["EPHB4",], main = "EPHB4", axisnames = FALSE, ylab = "count", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 2)
barplot(tgt@assays$RNA@data["VCAM1",], main = "VCAM1", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["IL1R1",], main = "IL1R1", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
barplot(tgt@assays$RNA@data["NR2F2",], main = "NR2F2", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
```


```{r}
saveRDS(CB_EC, "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/CB_SubAnnot.RDS")
```