---
title: "External SCEC Dataset 2"
author: "Bono"
date: "15/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After realizing the SCEC population does not contain enough cells, I will introduce two additional samples of SCEC derived from wildtype adult mouse (6-8 wks) for my comparison between brain and spinal cord ECs.

* Sample = GSM4955359


```{r, include=FALSE}
load_lib <- "Load Libraries"
```
## `r load_lib`
* Load all the necessary libraries

```{r load_lib, message=FALSE}
library(Seurat) 
library(dplyr)
library(scater)
library(SingleCellExperiment)
library(here)
library(ggplot2)
library(scran)
library(scuttle)
library(biomaRt)
```


#Read mouse spinal cord matrix dataset (GSM4955359)(Sample1)
```{r}
SCEC_2 <- read.csv(here("data/SC_external_dataset/GSM4955359_qc_filtered_feature_bc_matrix_uninj_sample1.txt"), sep = "\t", row.names = 1, header = T)
```


#Convert mouse gene names to human gene names
```{r}
ensembl <- useMart("ensembl")
ensembl_human <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
ensembl_mouse <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

mouse_genes <- rownames(SCEC_2)

matched_genes <- getLDS(attributes = "mgi_symbol", 
                        filters = "mgi_symbol",
                        values = mouse_genes,
                        mart = ensembl_mouse,
                        attributesL = "hgnc_symbol",
                        martL = ensembl_human)
```


#Save human gene nomenclature to mouse dataset
```{r}
matched_genes <- subset(matched_genes, matched_genes$HGNC.symbol != "")

subset_boul <- rownames(SCEC_2) %in% matched_genes$MGI.symbol

SCEC_2 <- SCEC_2[subset_boul,]

subset_boul_2 <- matched_genes$MGI.symbol %in% rownames(SCEC_2)

matched_genes <- matched_genes[match(rownames(SCEC_2), matched_genes$MGI.symbol),]

rownames(SCEC_2) <- make.unique(rownames(SCEC_2))

matched_genes$MGI.symbol <- make.unique(matched_genes$MGI.symbol)

matched_genes$HGNC.symbol <- make.unique(matched_genes$HGNC.symbol)

summary(rownames(SCEC_2) == matched_genes$MGI.symbol)

rownames(SCEC_2) <- matched_genes$HGNC.symbol
```


```{r}
barcode_meta <- read.table(here("data/SC_external_dataset/GSE162610_barcode_metadata.tsv"))

meta_1 <- barcode_meta[colnames(SCEC_2),]

SCEC_2 <- CreateSeuratObject(counts = SCEC_2, assay = "RNA", meta.data = meta_1)

SCEC_2.sce <- as.SingleCellExperiment(SCEC_2)
```


#Identify outliers of the mouse SC dataset using Scater
#Based on median absolute deviation (Scater)
```{r}
out_mito <- isOutlier(SCEC_2.sce$percent_mt, log = TRUE, type = "higher", nmads = 3)

out_feature <- isOutlier(SCEC_2.sce$nFeature_RNA, log = TRUE, type = "both", nmads = 3)

out_lib <- isOutlier(SCEC_2.sce$nCount_RNA, log = TRUE, type = "both", nmads = 3)

SCEC_2.sce <- SCEC_2.sce[, !c(out_mito | out_feature | out_lib)]

out_mito <- isOutlier(SCEC_2.sce$percent_mt, log = TRUE, type = "higher", nmads = 3)

out_feature <- isOutlier(SCEC_2.sce$nFeature_RNA, log = TRUE, type = "both", nmads = 3)

out_lib <- isOutlier(SCEC_2.sce$nCount_RNA, log = TRUE, type = "both", nmads = 3)

plot_mito <- plotColData(SCEC_2.sce, y = "percent_mt", colour_by = I(out_mito))

plot_feat <- plotColData(SCEC_2.sce, y = "nFeature_RNA", colour_by = I(out_feature))

plot_lib <- plotColData(SCEC_2.sce, y = "nCount_RNA", colour_by = I(out_feature))

gridExtra::grid.arrange(plot_mito, plot_feat, plot_lib, ncol = 3)
```


#Normalize mouse SC dataset using SCRAN from Bioconductor
#Build SeuratObject again containing raw counts (counts) and logcounts (data)
```{r}
clust_1 <- quickCluster(SCEC_2.sce)

SCEC_2.sce <- computeSumFactors(SCEC_2.sce, clusters = clust_1)

SCEC_2.sce <- logNormCounts(SCEC_2.sce)

SCEC_2 <- CreateSeuratObject(counts = counts(SCEC_2.sce), meta.data = as.data.frame(colData(SCEC_2.sce)))

SCEC_2 <- SetAssayData(object = SCEC_2, slot = "data", new.data = logcounts(SCEC_2.sce))
```


#Identify HVGs for dimRed and clustering
#2000 HVGs chosen here
```{r cluster_EC_Peri, message=FALSE}
#Calculate based on variance of log-expression of genes across cells. 2000 HVGs are selected.
SCEC_2 <- FindVariableFeatures(SCEC_2, selection.method = "vst", nfeatures = 2000)
top10_vg <- head(VariableFeatures(SCEC_2), 10)
top10_vg
```


#Plot HVGs from Seurat with and without labels
```{r, fig.width = 15, fig.height= 8, message= FALSE, warning=FALSE}
var_genes <- VariableFeaturePlot(SCEC_2)
labels <- LabelPoints(plot = var_genes, points = top10_vg, repel = TRUE)
var_genes + labels
```


#Perform PCA
```{r, message=FALSE}
all.genes <- rownames(SCEC_2)
SCEC_2 <- ScaleData(SCEC_2, features = all.genes) #save all variable genes 
```


```{r, message=FALSE}
#Perform PCA on log-normalised expression values of genes. Number of genes for PCA is based on the HVGs selected above.
SCEC_2 <- RunPCA(SCEC_2, features = VariableFeatures(object=SCEC_2))
```


```{r, fig.width= 10, fig.height= 8}
VizDimLoadings(SCEC_2, dims = 1:2, reduction = "pca") 
```


```{r}
DimPlot(SCEC_2, reduction = "pca")
```


```{r}
ElbowPlot(SCEC_2, ndims = 50)
```


```{r, include=FALSE}
cluster <- "Clustering Analysis"
```
# `r cluster`


#The mouse dataset has been previously annotated. I will check whether the annotations are accurate
#Cluster resolution set to 0.5
```{r, message=FALSE, warning= FALSE}
Idents(SCEC_2) <- "celltype"
SCEC_2 <- FindNeighbors(SCEC_2, dims = 1:30) #KNN
SCEC_2 <- FindClusters(SCEC_2, resolution = 0.5) #Louvain algorithm
Idents(SCEC_2) <- "RNA_snn_res.0.5"
```


```{r, message=FALSE, warning= FALSE}
##Plot UMAP and t-SNE
SCEC_2 <- RunUMAP(SCEC_2, dims = 1:30)
SCEC_2 <- RunTSNE(SCEC_2, dims = 1:30)

DimPlot(SCEC_2, reduction = "umap", label = T)
DimPlot(SCEC_2, reduction = "tsne", label = T)
DimPlot(SCEC_2, reduction = "pca", label = F, pt.size = 2)

Idents(SCEC_2) <- "celltype"
Endo <- WhichCells(SCEC_2, idents = "Endothelial")

DimPlot(SCEC_2, group.by = "celltype", cells.highlight = Endo) + scale_color_manual(labels = c("Other", "Endothelial"), values = c("grey", "darkblue")) + ggtitle("GSM4955359")

DimPlot(SCEC_2, group.by = "celltype")
```


```{r, include=FALSE}
feature_plt_1 <- "Feature Plots (EC genes)"
```
## `r feature_plt_1`


```{r feature_plt_1, fig.width=15, fig.height=13}
#EC  Biomarkers
FeaturePlot(SCEC_2, features = c("FLT1",
                              "CLDN5",
                              "ABCG2",
                              "ELOVL7",
                              "PECAM1",
                              "MECOM"
                              ))
```


```{r, include=FALSE}
feature_plt_2 <- "Feature Plots (Pericyte genes)"
```
## `r feature_plt_2`


```{r feature_plt_2, fig.width=15, fig.height=12}

#Pericyte Biomarkers
FeaturePlot(SCEC_2, features = c("PDGFRB",
                              "NOTCH3",
                              "PRKG1",
                              "CSPG4",
                              "GPC5",
                              "RGS5",
                              "PDE7B"
                              ))
```


#Save SCEC sample 1
```{r, message= FALSE}
SCEC_2 <- subset(SCEC_2, subset=celltype=="Endothelial")
```


```{r, include=FALSE}
feature_plt_3 <- "Verify subsetted SCEC population"
```
## `r feature_plt_3`


```{r}
FeaturePlot(SCEC_2, feature = c("CLDN5", "FLT1", "PECAM1", "ABCB1")) #EC marker

FeaturePlot(SCEC_2, feature = c("PDGFRB", "CSPG4", "GPC5", "RGS5")) #Pericyte marker
```

```{r}
saveRDS(SCEC_2, "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/SCEC_mouse_2.RDS")
```