z---
title: "Merge SCEC population"
author: "Bono"
date: "16/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is written to integrate the mouse external SCEC dataset with Seeker BA4 & CB & CSC endothelial population. The integrated product of this step will be used for clustering, DE testing, as well as identifying differences between brain and spinal cord ECs.

* External mouse SCEC dataset is a young adult of 8-10 weeks old
* GSE162610
* QC-ed through Scater
* Normalized through Scran
* EC subtype annotated either Arterial, Capillary, or Vein


```{r, include=FALSE}
load_lib <- "Load Libraries"
```
## `r load_lib`
* Load all the necessary libraries

```{r load_lib, message=FALSE}
library(Seurat) 
library(dplyr)
library(scater)
library(SingleCellExperiment)
library(here)
library(ggplot2)
library(scran)
library(scuttle)
library(Matrix)
library(biomaRt)
library(reshape2)
```


#Import the integrated and downsampled GSE162610 dataset
#The two mouse samples had been previously QCed and normalised
#Can integrate with Seeker data directly
```{r}
SCEC_mouse <- readRDS(here("data/mouse_downsampled.RDS"))
```


#Pre-processing of Seeker data prior CCA integration 
```{r, message= FALSE}
merged_EC <- readRDS(here("data/merged_EC.RDS"))

merged_EC <- FindVariableFeatures(merged_EC, selection.method = "vst", nfeatures = 2000) #Find Top 2000 HVGs
```


#Find anchors using Seeker and mouse SCEC data
```{r}
features <- SelectIntegrationFeatures(object.list = list(merged_EC, SCEC_mouse))

#merged_EC <- ScaleData(merged_EC, features = features, verbose = FALSE)
#merged_EC <- RunPCA(merged_EC, features = features, verbose = FALSE)

#SCEC_mouse <- ScaleData(SCEC_mouse, features = features, verbose = FALSE)
#SCEC_mouse <- RunPCA(SCEC_mouse, features = features, verbose = FALSE)

ec_anchors <- FindIntegrationAnchors(object.list = list(merged_EC, SCEC_mouse), anchor.features = features, normalization.method = "LogNormalize", reduction = "cca", k.filter = 200, k.anchor = 5)
```


#Perform CCA integration
#Change default assay of integrated object to "integrate"
```{r}
ECs_comb <- IntegrateData(anchorset = ec_anchors, k.weight = 100)

DefaultAssay(ECs_comb) <- "integrated"
```



#Perform an integrated analysis
#Perform PCA of integrated objet
```{r}
all_genes <- rownames(ECs_comb)
        
ECs_comb <- ScaleData(ECs_comb, features = all_genes)

ECs_comb <- RunPCA(ECs_comb, npcs = 40, verbose = FALSE)
```


#Perform clustering of integrated object
```{r}
ECs_comb <- RunUMAP(ECs_comb, reduction = "pca", dims = 1:20)
ECs_comb <- FindNeighbors(ECs_comb, reduction = "pca", dims = 1:20)
ECs_comb <- FindClusters(ECs_comb, resolution = 0.45)
```


#Visualize clustering of integrated object
#Cluster 2 = Mainly arterial cells 
#Cluster 5 = Mainly vein-like SCECs
```{r}
plot_sample <- DimPlot(ECs_comb, reduction = "umap", group.by = "sample")
plot_sample + ggtitle("Sample")

DimPlot(ECs_comb, reduction = "umap", group.by = "Tissues") + ggtitle("BEC + SCEC") + theme(plot.title = element_text(hjust = 0.5))

plot_clust <- DimPlot(ECs_comb, reduction = "umap", label = TRUE)
plot_clust + ggtitle("Sub-clustering") + theme(plot.title = element_text(hjust=0.5))

plot_subtype <- DimPlot(ECs_comb, reduction = "umap", group.by = "subtype")
plot_subtype + ggtitle("Endothelial Subtype")

plot_ind <- DimPlot(ECs_comb, reduction = "umap", group.by = "BBN")
plot_ind + ggtitle("Individual")
```


#Number of cells per cluster
```{r, echo = TRUE}
CellsPerCluster <- as.data.frame(tapply(
    ECs_comb@meta.data$orig.ident, ECs_comb@meta.data$integrated_snn_res.0.45, function(x) length(x) ))

names(CellsPerCluster) <- "NumberOfCells"

CellsPerCluster$RNA_snn_res.0.5 <- rownames(CellsPerCluster)

CellsPerCluster$RNA_snn_res.0.5 <- rownames(CellsPerCluster)

CellsPerCluster
```


#Set default assay back to RNA
#Save integrated object
```{r}
DefaultAssay(ECs_comb) <- "RNA"

saveRDS(ECs_comb, file = "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/int_ECs.RDS")
```
