---
title: "GSE135893 (Public lung dataset)"
author: "B180131 dissertation"
date: "30/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the scRNA-seq analysis workflow for comparing human brain endothelial cells vs human lung endothelial cells. This allows for better understanding of how tissue-specific ECs provide specific requirements for a given vascular bed. Required packages involve Seurat and Bioconductor. 

Lung dataset was obtained from GSE135893 and had been annotated with all cell types including endothelial cells. This script aims to carry out QC on the dataset.


#Load libraries
```{r load_lib, echo=FALSE}
library(Seurat) 
library(dplyr)
library(scater)
library(SingleCellExperiment)
library(here)
library(ggplot2)
library(ggsci)
library(scales)
library(scran)
```


Pick colour paletts

```{r, echo= FALSE}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)
mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)
show_col(mycoloursP, labels =F)
```


#Load data
#Since the compute for this block takes a lot of time (file too large)
#This has already been ran and we will import the RDS object immediately from the block below
```{r, echo = FALSE}
#data_dir <- "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/GSE135893/"

#GSE135893 <- Read10X(data.dir = data_dir, gene.column = 1)

#barcode_meta <- read.csv(here("data/GSE135893/GSE135893_IPF_metadata.csv"), row.names = 1)

#meta_1 <- barcode_meta[colnames(GSE135893),]

#seur_obj <- CreateSeuratObject(counts=GSE135893, meta.data = meta_1)

#saveRDS(seur_obj, file = "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/full_matrix.RDS")
```


#Import seurat object of entire dataset 
```{r}
seur_obj <- readRDS(here("data/GSE135893/full_matrix.RDS"))
```


#Subset only control samples for pre-processing
GSE135893 contains 8 control lung biospies with identified age. We excluded two control samples because these captured very little lung-ECs (< 40). The control samples to be used are:

* VUHD69 
* VUHD65
* VUHD67
* VUHD68
* VUHD70
* VUHD071

#Subset 

* Subset only control samples from the raw dataset
```{r}
cont_samp <- subset(seur_obj, subset=(Sample_Name=="VUHD69" | Sample_Name=="VUHD071" | Sample_Name=="VUHD65" | Sample_Name=="VUHD67" | Sample_Name=="VUHD68" | Sample_Name=="VUHD70"))
```


#Quality control of control dataset (Scater)
Remove cells based on outlier as 3 MAD

```{r}
cont_samp.sce <- as.SingleCellExperiment(cont_samp) 

out_mito <- isOutlier(cont_samp.sce$percent.mt, log = TRUE, type = "higher", nmads = 3)

out_feat <- isOutlier(cont_samp.sce$nFeature_RNA, log = TRUE, type = "both", nmads = 3)

out_count <- isOutlier(cont_samp.sce$nCount_RNA, log = TRUE, type = "both", nmads = 3)

discard <- out_mito | out_feat | out_count

cont_samp.sce <- cont_samp.sce[, !discard] 
```


#Normalize dataset
Normalise using Scran to stay consistent with Luise's dataset

```{r}
clusters <- quickCluster(cont_samp.sce)

cont_samp.sce <- computeSumFactors(cont_samp.sce, clusters = clusters)

cont_samp.sce <- logNormCounts(cont_samp.sce)
```


#Convert to SeuratObject
```{r}
cont_samp <- CreateSeuratObject(counts = counts(cont_samp.sce), meta.data = as.data.frame(colData(cont_samp.sce)))

cont_samp <- SetAssayData(object = cont_samp, slot = "data", new.data = logcounts(cont_samp.sce))
```


#Save cont_samp object
```{r}
saveRDS(cont_samp, file = "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/cont_samp.RDS")
```


#Sample integration on the control dataset
```{r}
cont.list <- SplitObject(cont_samp, split.by = "Sample_Name")

for (i in 1:length(cont.list)){
        cont.list[[i]] <- NormalizeData(cont.list[[i]], verbose = FALSE)
        cont.list[[i]] <- FindVariableFeatures(cont.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
```


```{r}
features <- SelectIntegrationFeatures(object.list = cont.list, verbose = FALSE)

cont.anchors <- FindIntegrationAnchors(object.list = cont.list, anchor.features = features, dims = 1:30, verbose = FALSE)

cont.int <- IntegrateData(anchorset= cont.anchors, dims = 1:30, verbose = FALSE)
```


#UMAP visualization of integrated object
```{r}
DefaultAssay(cont.int) <- "integrated"

all_genes <- rownames(cont.int)

cont.int <- ScaleData(cont.int, verbose = FALSE, block.size = 150) 

cont.int <- RunPCA(cont.int, npcs = 30, verbose =FALSE)

cont.int <- RunUMAP(cont.int, reduction = "pca", dims = 1:30)
```


#Visualize
```{r}
DimPlot(cont.int, reduction = "umap", group.by = "celltype", cols = mycoloursP[20:50])

DimPlot(cont.int, reduction = "umap", group.by = "Sample_Name")

DefaultAssay(cont.int) <- "RNA"
FeaturePlot(cont.int, features = c("CLDN5", "PECAM1"))
```


#Save integrated lung dataset
````{r}
DefaultAssay(cont.int) <- "integrated"

saveRDS(cont.int, file = "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/int_lung.RDS")
```