---
title: "Subtype annotaiton of pulmonary ECs"
author: "Bono"
date: "03/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load_lib, message=FALSE}
library(Seurat) 
library(dplyr)
library(scater)
library(SingleCellExperiment)
library(here)
library(ggplot2)
library(scran)
library(fgsea)
library(msigdbr)
library(clusterProfiler)
library(GSEABase)
library(org.Hs.eg.db)
library(SCINA)
library(scSorter)
library(cowplot)
```


# Following previous script, the endothelial dataset from GSE135893 has been integrated to get rid of batch effect

* Here, we will annotate the subtype of ECs for this integrated object using scSorter
* The UMAP visualization of the pulmonary-ECs generated three distinct subclusters. It is believed that these three subclusters were separated based on subtype
* Subtype markers are primarily referenced from Haln's research but also from literature


```{r}
comb_lungEC <- readRDS("/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/lungEC_int.RDS")

DefaultAssay(comb_lungEC) <- "RNA"
```


#Generate marker file as input for scSorter

* Arterial markers: GJA5 (Su), CXCL12 (Haln), BMX (Valendewijck)
* Capillary markers: CA4 (Haln & Fleming et al), PRX (Wang), RGCC (Kalucka)
* Vein markers: ACKR1 (Thiriot), NR2F2 (Schupp), CPE (Schupp)



#Feature plot for the visualization of pulmonary-EC subtype markers 

*Arterial markers
```{r}
plot_art <- FeaturePlot(comb_lungEC, features = c("GJA5", "CXCL12", "BMX"))
plot_grid(ggdraw() + draw_label("Pulmonary arterial EC markers", size = 13, fontface = "bold"), plot_art, ncol = 1, rel_heights = c(0.1,1))
```


*Capillary markers
```{r}
plot_cap <- FeaturePlot(comb_lungEC, features = c("CA4", "RGCC", "PRX"))
plot_grid(ggdraw() + draw_label("Pulmonary capillary EC markers", size = 13, fontface = "bold"), plot_cap, ncol = 1, rel_heights = c(0.1,1))
```


*Vein markers
```{r}
plot_vein <- FeaturePlot(comb_lungEC, features = c("ACKR1", "NR2F2", "CPE"))
plot_grid(ggdraw() + draw_label("Pulmonary vein EC markers", size = 13, fontface = "bold"), plot_vein, ncol = 1, rel_heights = c(0.1,1))
```

#scSorter uses the data slot from SeuratObject
```{r}
DefaultAssay(comb_lungEC) <- "RNA"
```


#Generate marker file
```{r}
Type <- c("Capillary", "Capillary", "Capillary", "Arterial", "Arterial", "Arterial", "Vein", "Vein", "Vein")

Marker <- c("CA4", "RGCC", "PRX", "GJA5", "CXCL12", "BMX", "ACKR1", "CPE", "NR2F2")

Weight <- c(5, 5, 2, 2, 5, 5, 5, 3, 5)

Markers <- cbind(Type, Marker, Weight)

Markers <- as.data.frame(Markers)
```


#Subtype annotate
```{r}
annotate <- scSorter(comb_lungEC@assays$RNA@data, Markers, alpha = 0)
```


#Add to metadata
```{r}
comb_lungEC$subtype <- annotate$Pred_Type

Idents(comb_lungEC) <- "subtype"

comb_lungEC <- subset(comb_lungEC, subset = (subtype == "Arterial" | subtype == "Capillary" | subtype == "Vein"))
```


#Visualize subtype annotation
```{r}
DimPlot(comb_lungEC, reduction = "umap", group.by = "subtype", cols = c("coral2", "chartreuse4","cornflowerblue")) + ggtitle("Pulmonary EC subtype") + theme(plot.title = element_text(hjust = 0.5))
```


#Save pre-processed pulmonary-EC dataset
```{r}
saveRDS(comb_lungEC, file = "/Users/bonocheong/Desktop/Research_Project/Datasets/Dissertation/data/pEC_SubAnnot.RDS")
```


#Number of cells in each subtype
```{r}
art <- comb_lungEC[,comb_lungEC@meta.data$subtype == "Arterial"]

capi <- comb_lungEC[,comb_lungEC@meta.data$subtype == "Capillary"]

veins <- comb_lungEC[,comb_lungEC@meta.data$subtype == "Vein"]

tgt <- merge(art, c(capi, veins), add.cell.ids = c("Arterial", "Capillary", "Vein"))
```


```{r, echo = TRUE}
CellsPerSubtype <- as.data.frame(tapply(
    comb_lungEC@meta.data$orig.ident, comb_lungEC@meta.data$subtype, function(x) length(x) ))

names(CellsPerSubtype) <- "NumberOfCells"

CellsPerSubtype$Subtype <- rownames(CellsPerSubtype)

CellsPerSubtype$Subtype <- rownames(CellsPerSubtype)

CellsPerSubtype
```


```{r}
col_art <- rep("coral2",ncol(tgt[,tgt@meta.data$subtype=="Arterial"]))

col_capi <- rep("chartreuse4",ncol(tgt[,tgt@meta.data$subtype=="Capillary"]))

col_vein <- rep("cornflowerblue",ncol(tgt[,tgt@meta.data$subtype=="Vein"]))

palette <- c(col_art, col_capi, col_vein)

tgt$color <- palette
```


#Generate expression barplot
```{r}
par(mfrow = c(3, 1))
barplot(tgt@assays$RNA@data["GJA5",], main = "GJA5", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5) 

barplot(tgt@assays$RNA@data["CXCL12",], main = "CXCL12", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["BMX",], main = "BMX", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["CA4",], main = "CA4", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["RGCC",], main = "RGCC", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["PRX",], main = "PRX", axisnames = FALSE, border = tgt$color, ylab = "Expression", cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["ACKR1",], main = "ACKR1", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["CPE",], main = "CPE", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)

barplot(tgt@assays$RNA@data["NR2F2",], main = "NR2F2", axisnames = FALSE, ylab = "Expression", border = tgt$color, cex.axis = 2, cex.main = 3, cex.lab = 1.5)
```