---
title: "05_tissue_DE_testing"
author: "B180131"
date: "07/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# From previous script, BECs and PECs were integrated into a single SeuratObject for analysis

* It is found that these two tissue-specific ECs do not overlap and the same outcome has appeared in previous literature (Feng et al 2019)

# From this we will proceed to a direct DE testing between BECs and PECs:
* Capillary-BECs vs Capillary-PECs
* Vein-BECs vs Vein-PECs
* Arterial-BECs vs Arterial-BECs

# It is worth also just performing a global DE testing
* BECs vs PECs (regardless of subtype)


```{r load_lib, message=FALSE}
library(Seurat) 
library(dplyr)
library(scater)
library(SingleCellExperiment)
library(here)
library(ggplot2)
library(scran)
library(stringr)
library(gtools)
library(scSorter)
library(MAST)
library(VennDetail)
library(gridExtra)
library(grid)
library(fgsea)
library(msigdbr)
library(clusterProfiler)
library(GSEABase)
library(ensembldb)
library(org.Hs.eg.db)
library(enrichplot)
library(ggsci)
library(scales)
```


Pick colour paletts

```{r, echo= FALSE}
mypal <- pal_npg("nrc", alpha = 0.7)(10)
mypal2 <-pal_tron("legacy", alpha = 0.7)(7)
mypal3 <- pal_lancet("lanonc", alpha = 0.7)(9)
mypal4 <- pal_simpsons(palette = c("springfield"), alpha = 0.7)(16)
mypal5 <- pal_rickandmorty(palette = c("schwifty"), alpha = 0.7)(6)
mypal6 <- pal_futurama(palette = c("planetexpress"), alpha = 0.7)(5)
mypal7 <- pal_startrek(palette = c("uniform"), alpha = 0.7)(5)
mycoloursP<- c(mypal, mypal2, mypal3, mypal4, mypal5, mypal6, mypal7)
show_col(mycoloursP, labels =F)
```


#Load integrated BEC and PEC object
```{r}
EC.comb <- readRDS(here("data/BEC_PEC.RDS"))

DefaultAssay(EC.comb) <- "RNA"

Idents(EC.comb) <- "sub_tis"
```


#Direct comparison of BEC and PEC according to subtype

* Capillary BEC vs Capillary PEC
* Arterial BEC vs Arterial PEC
* Vein BEC vs Vein PEC


```{r}
DEG_capi <- FindMarkers(EC.comb, ident.1 = "Capillary-BEC", ident.2 = "Capillary-PEC", test.use = "MAST", logfc.threshold = 1, min.pct = 0.5, verbose = FALSE)

DEG_capi_up <- DEG_capi[DEG_capi$p_val_adj < 0.05 & DEG_capi$avg_log2FC > 1 & (DEG_capi$pct.1 - DEG_capi$pct.2) > 0.5,]

DEG_capi_up

DEG_capi_down <- DEG_capi[DEG_capi$p_val_adj < 0.05 & DEG_capi$avg_log2FC < -1 & (DEG_capi$pct.1 - DEG_capi$pct.2) < -0.5,]

DEG_capi_down
```


```{r}
DEG_art <- FindMarkers(EC.comb, ident.1 = "Arterial-BEC", ident.2 = "Arterial-PEC", test.use = "MAST", logfc.threshold = 1, min.pct = 0.5, verbose = FALSE)

DEG_art_up <- DEG_art[DEG_art$p_val_adj < 0.05 & DEG_art$avg_log2FC > 1 & (DEG_art$pct.1 - DEG_art$pct.2) > 0.5,]

DEG_art_up

DEG_art_down <- DEG_art[DEG_art$p_val_adj < 0.05 & DEG_art$avg_log2FC < -1 & (DEG_art$pct.1 - DEG_art$pct.2) < -0.5,]

DEG_art_down
```


```{r}
DEG_vein <- FindMarkers(EC.comb, ident.1 = "Vein-BEC", ident.2 = "Vein-PEC", test.use = "MAST", logfc.threshold = 1, min.pct = 0.5, verbose = FALSE)

DEG_vein_up <- DEG_vein[DEG_vein$p_val_adj < 0.05 & DEG_vein$avg_log2FC > 1 & (DEG_vein$pct.1 - DEG_vein$pct.2) > 0.5,]

DEG_vein_up

DEG_vein_down <- DEG_vein[DEG_vein$p_val_adj < 0.05 & DEG_vein$avg_log2FC < -1 & (DEG_vein$pct.1 - DEG_vein$pct.2) < -0.5,]

DEG_vein_down
```


```{r}
all_DEGs <- rbind(DEG_capi_up, DEG_art_up, DEG_vein_up, DEG_capi_down, DEG_art_down, DEG_vein_down)
```


#Plot Heatmap
```{r}
my_levels <- c("Capillary-BEC", "Arterial-BEC", "Vein-BEC", "Capillary-PEC", "Arterial-PEC", "Vein-PEC")
EC.comb$sub_tis <- factor(x = EC.comb$sub_tis, levels = my_levels)

Idents(EC.comb) <- "sub_tis"
        
EC.comb <- ScaleData(EC.comb, features = rownames(EC.comb), verbose = FALSE)

ec.downsample <- subset(EC.comb, downsample= 50)
        
heatmap <- DoHeatmap(EC.comb, features = rownames(all_DEGs), label = TRUE, cells = Cells(ec.downsample), size = 3) + scale_fill_viridis_b(option = "viridis")

heatmap + theme(axis.text.y = element_text(size = 0)) 
```


# Generate DEG table

Generate the top 10 DEGs for each subtype between BEC and PEC
```{r}
deg_cap_top10 <- DEG_capi_up[1:10,]
deg_cap_down10 <- DEG_capi_down[1:10,]
capillary_degs <- c(rownames(deg_cap_top10), rownames(deg_cap_down10))

deg_art_top10 <- DEG_art_up[1:10,]
deg_art_down10 <- DEG_art_down[1:10,]
arterial_degs <- c(rownames(deg_art_top10), rownames(deg_art_down10))

deg_vein_top10 <- DEG_vein_up[1:10,]
deg_vein_down10 <- DEG_vein_down[c(1,2,3,4,5,7,11,12,14,15),]
vein_degs <- c(rownames(deg_vein_top10), rownames(deg_vein_down10))

deg.table <- data.frame(capillary_degs, arterial_degs, vein_degs)
colnames(deg.table) <- c("Capillary", "Arterial", "Vein")

deg.table
deg.table <- as.matrix(deg.table)
```


#Output DEG table
```{r}
t1 <- ttheme_default(
        #Set color for table holding DEGs
        core=list(
        bg_params = list(fill=c(rep("peachpuff3",
                                    length.out=10), rep("slategray3", length.out=10)),
                         col = c(rep("peachpuff3",
                                    length.out=10), rep("slategray3", length.out=10)))),
        #Change column header to 
        colhead = list(fg_params=list(col="white"),
                       bg_params=list(fill = "dimgrey", col = "dimgrey"))
        ) 


grid.table(deg.table[,1:3], theme = t1)
```


#####PERFORM GSEA####


# GO analysis of brain markers for each subtype
```{r}
hs <- org.Hs.eg.db
```


```{r}
deg_capi_brain <- DEG_capi_up$avg_log2FC

names(deg_capi_brain) <- toupper(rownames(DEG_capi_up))

deg_capi_brain <- sort(deg_capi_brain, decreasing = T)

deg_capi_brain_entrez <- select(hs, keys = names(deg_capi_brain), columns = "ENTREZID", keytype =  "SYMBOL")
```


#GO analysis on capillary DEGs
```{r}
deg_capi_kk <- enrichGO(gene =deg_capi_brain_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_capi_kk <- setReadable(deg_capi_kk, OrgDb = hs, keyType = "ENTREZID")

deg_capi_readable <- as.data.frame(deg_capi_kk)

barplot(deg_capi_kk, showCategory = 10)

#Show GO result of the terms of interest and are significant after multiple correction
barplot(deg_capi_kk, showCategory = deg_capi_kk@result$Description[c(3,5,7,10,21)], color = "qvalue") + ggtitle("Capillary Brain EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + xlab("Gene count")
```


#GO analysis on arterial DEGs
```{r}
deg_art_brain <- DEG_art_up$avg_log2FC

names(deg_art_brain) <- toupper(rownames(DEG_art_up))

deg_art_brain <- sort(deg_art_brain, decreasing = T)

deg_art_brain_entrez <- select(hs, keys = names(deg_art_brain), columns = "ENTREZID", keytype =  "SYMBOL")
```


```{r}
deg_art_kk <- enrichGO(gene =deg_art_brain_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_art_kk <- setReadable(deg_art_kk, OrgDb = hs, keyType = "ENTREZID")

deg_art_readable <- as.data.frame(deg_art_kk)

barplot(deg_art_kk, showCategory = 10)

barplot(deg_art_kk, showCategory = deg_art_kk@result$Description[c(2, 3, 4, 5, 8, 17)], color = "qvalue") + ggtitle("Arterial Brain EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + xlab("Gene count")

art_GO_of_interest <- dplyr::filter(deg_art_kk, p.adjust > 0.0014985, p.adjust < 0.014508364)
```


#GO on vein BECs
```{r}
deg_vein_brain <- DEG_vein_up$avg_log2FC

names(deg_vein_brain) <- toupper(rownames(DEG_vein_up))

deg_vein_brain <- sort(deg_vein_brain, decreasing = T)

deg_vein_brain_entrez <- select(hs, keys = names(deg_vein_brain), columns = "ENTREZID", keytype =  "SYMBOL")
```


```{r}
deg_vein_kk <- enrichGO(gene =deg_vein_brain_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_vein_kk <- setReadable(deg_vein_kk, OrgDb = hs, keyType = "ENTREZID")

deg_vein_readable <- as.data.frame(deg_vein_kk)

barplot(deg_vein_kk, showCategory = deg_vein_kk@result$Description[c(1, 2, 3, 5, 6)], color = "qvalue") + ggtitle("Vein Brain EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + xlab("Gene count")

vein_GO_of_interest <- dplyr::filter(deg_vein_kk, p.adjust < 0.035)
```



# From the GO analysis, we found that there is a global endothelial upregulation of various molecule transport genes

* Generate violin plot to see how these genes are upregulated in BEC vs PEC

* Xenobiotic transport

```{r}
abcb1 <- VlnPlot(EC.comb, features = "ABCB1") & scale_fill_manual(values = mycoloursP[c(1,6,15,16,20,22)]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

abcg2 <- VlnPlot(EC.comb, features = "ABCG2") & scale_fill_manual(values = mycoloursP[c(1,6,15,16,20,22)]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

slc2a1 <- VlnPlot(EC.comb, features = "SLC2A1") & scale_fill_manual(values = mycoloursP[c(1,6,15,16,20,22)]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

slc7a5 <- VlnPlot(EC.comb, features = "SLC7A5") & scale_fill_manual(values = mycoloursP[c(1,6,15,16,20,22)]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

gridExtra::grid.arrange(abcb1, abcg2, slc2a1, slc7a5, nrow = 4) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()
```


* Amino acid transport

```{r}
slc7a1 <- VlnPlot(EC.comb, features = "SLC7A1") & scale_fill_manual(values = mycoloursP[20:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

slc7a11 <- VlnPlot(EC.comb, features = "SLC7A11") & scale_fill_manual(values = mycoloursP[20:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

slc1a1 <- VlnPlot(EC.comb, features = "SLC1A1") & scale_fill_manual(values = mycoloursP[20:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

gridExtra::grid.arrange(slc7a1, slc7a11, slc1a1, nrow = 4)
```


* Lipid transport

```{r}
atp10a <- VlnPlot(EC.comb, features = "ATP10A") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

ncoa1 <- VlnPlot(EC.comb, features = "NCOA1") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

esyt2 <- VlnPlot(EC.comb, features = "ESYT2") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

gridExtra::grid.arrange(atp10a, ncoa1, esyt2, nrow = 4)
```


##########################################################################

#Now, gene ontology is performed on the lung-EC markers

#Capillary lung marker GO
```{r}
deg_capi_lung <- DEG_capi_down$avg_log2FC

names(deg_capi_lung) <- toupper(rownames(DEG_capi_down))

deg_capi_lung <- sort(deg_capi_lung, decreasing = T)

deg_capi_lung_entrez <- select(hs, keys = names(deg_capi_lung), columns = "ENTREZID", keytype =  "SYMBOL")
```


```{r}
deg_capi_kk2 <- enrichGO(gene =deg_capi_lung_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_capi_kk2 <- setReadable(deg_capi_kk2, OrgDb = hs, keyType = "ENTREZID")

deg_capi_readable2 <- as.data.frame(deg_capi_kk2)

barplot(deg_capi_kk2, showCategory = 10)

#Show GO result of the terms of interest and are significant after multiple correction
barplot(deg_capi_kk2, x = "count", showCategory = deg_capi_kk2@result$Description[c(17, 29, 30, 33, 34)], color = "qvalue") + ggtitle("Capillary Lung EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + scale_fill_viridis_b(option = "viridis") + xlab("Gene count")
```


#Arterial lung marker GO
```{r}
deg_art_lung <- DEG_art_down$avg_log2FC

names(deg_art_lung) <- toupper(rownames(DEG_art_down))

deg_art_lung <- sort(deg_art_lung, decreasing = T)

deg_art_lung_entrez <- select(hs, keys = names(deg_art_lung), columns = "ENTREZID", keytype =  "SYMBOL")
```


```{r}
deg_art_kk2 <- enrichGO(gene =deg_art_lung_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_art_kk2 <- setReadable(deg_art_kk2, OrgDb = hs, keyType = "ENTREZID")

deg_art_readable2 <- as.data.frame(deg_art_kk2)

barplot(deg_art_kk2, showCategory = 10)

#Show GO result of the terms of interest and are significant after multiple correction
barplot(deg_art_kk2, x = "count", showCategory = deg_art_kk2@result$Description[c(22, 28, 31, 43, 44)], color = "qvalue") + ggtitle("Arterial Lung EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + scale_fill_viridis_b(option = "viridis") + xlab("Gene count")
```


#Vein lung marker GO
```{r}
deg_vein_lung <- DEG_vein_down$avg_log2FC

names(deg_vein_lung) <- toupper(rownames(DEG_vein_down))

deg_vein_lung <- sort(deg_vein_lung, decreasing = T)

deg_vein_lung_entrez <- select(hs, keys = names(deg_vein_lung), columns = "ENTREZID", keytype =  "SYMBOL")
```


```{r}
deg_vein_kk2 <- enrichGO(gene =deg_vein_lung_entrez$ENTREZID, OrgDb = hs, pvalueCutoff = 0.05, ont = "BP")

deg_vein_kk2 <- setReadable(deg_vein_kk2, OrgDb = hs, keyType = "ENTREZID")

deg_vein_readable2 <- as.data.frame(deg_vein_kk2)

barplot(deg_vein_kk2, showCategory = 10)

#Show GO result of the terms of interest and are significant after multiple correction
barplot(deg_vein_kk2, x = "count", showCategory = deg_vein_kk2@result$Description[c(27, 31, 32, 62,63)], color = "qvalue") + ggtitle("Vein Lung EC signature") + theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.y.left = element_text(size = 8.5)) + scale_fill_viridis_b(option = "viridis") + xlab("Gene count")
```


# From the GO analysis, we found that there is a global endothelial upregulation of viral response genes

* Generate violin plot to see how these genes are upregulated in PEC vs BEC

* Negative regulation of viral entry into host cell
```{r}
hla_drb1 <- VlnPlot(EC.comb, features = "HLA-DRB1") & scale_fill_manual(values = mycoloursP[c(1,2,3,8,16,21 )]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

ifitm2 <- VlnPlot(EC.comb, features = "IFITM2") & scale_fill_manual(values = mycoloursP[c(1,2,3,8,16,21 )]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

fcn3 <- VlnPlot(EC.comb, features = "FCN3") & scale_fill_manual(values = mycoloursP[c(1,2,3,8,16,21 )]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

ifitm3 <- VlnPlot(EC.comb, features = "IFITM3") & scale_fill_manual(values = mycoloursP[c(1,2,3,8,16,21 )]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()


gridExtra::grid.arrange(hla_drb1, ifitm2, fcn3, ifitm3, nrow = 4)
```


* Negative regulation of viral processes

```{r}
clta <- VlnPlot(EC.comb, features = "CLTA") & scale_fill_manual(values = mycoloursP[24:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

ctsd <- VlnPlot(EC.comb, features = "CTSD") & scale_fill_manual(values = mycoloursP[24:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

hla_drb5 <- VlnPlot(EC.comb, features = "HLA-DRB5") & scale_fill_manual(values = mycoloursP[24:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

hla_b <- VlnPlot(EC.comb, features = "HLA-B") & scale_fill_manual(values = mycoloursP[24:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

gridExtra::grid.arrange(clta, ctsd, hla_drb5, hla_b, nrow = 4)
```


* Antigen processing and presentation of exogenous antigen

EEF1A1/S100A11/NME2/ATP6V0C/PPIA

```{r}
eef1a1 <- VlnPlot(EC.comb, features = "EEF1A1") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

s100a11 <- VlnPlot(EC.comb, features = "S100A11") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

nme2 <- VlnPlot(EC.comb, features = "NME2") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

atp6v0c <- VlnPlot(EC.comb, features = "ATP6V0C") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

ppia <- VlnPlot(EC.comb, features = "PPIA") & scale_fill_manual(values = mycoloursP[43:60]) & theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank(), plot.title = element_blank(), axis.text.x.bottom = element_blank()) & NoLegend()

gridExtra::grid.arrange(eef1a1, s100a11, ppia, nrow = 4)
```


#Control for expression between BEC and PEC
```{r}
cldn5 <- VlnPlot(EC.comb, features = "CLDN5", pt.size = 0) & NoLegend()

pecam1 <- VlnPlot(EC.comb, features = "PECAM1", pt.size = 0) & NoLegend()

vwf <- VlnPlot(EC.comb, features = "VWF", pt.size = 0) & NoLegend()

gridExtra::grid.arrange(cldn5, pecam1, vwf, nrow = 3)
```